# Reproducible Research: Peer Assessment 1
## User: egrossman
Welcome! Here is my first project using RMarkdown.

I will be showing you how I analyze and transform data. For this specific data set, I am using data that is a result of how many steps a subject has taken. This data is taken in 5 second intervals across multiple days. Each row contains:

* The **date** this data was taken
* The Five second **interval** that the number of steps were recorded
* The number of **steps** that were taken in this interval

## Loading and preprocessing the data
This section is pretty stright-forward. I've loaded the dataset into R using the read.csv command. I've also defined the classes of each column.

```{r}
activityData <- read.csv('activity.csv',colClasses=c('integer','Date','integer'))
```


## What is mean total number of steps taken per day?
In order to generate the number of steps per day, I had to first take the sum of the **steps** for each **date**. The aggregate function does this nicely.

I've also  color coded the histogram as blue and made 10 break points for the frequency.

```{r}
stepsPerDay<-aggregate(steps ~date ,activityData, sum)
hist(x=stepsPerDay[,"steps"],main='Frequency of Steps',xlab='Steps per day',col='BLUE',breaks=10)
```

## What is the average daily activity pattern?

```{r}
averageSteps<-aggregate(steps ~interval ,activityData, mean)
## Get max value
maxSteps <- max(averageSteps$steps)
## Get interval associated with max
maxInterval<-averageSteps[which(averageSteps$steps == maxSteps),'interval']

## Plot data
with(data=averageSteps,plot(interval,steps,type='l',xlab='Intervals(seconds)'))
points(maxInterval,maxSteps,pch=16,col='RED',lwd=20)
text(x=maxInterval+220,y=maxSteps,labels=paste('Max Value    (',maxInterval,',',maxSteps,')'),cex=0.8)
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
